---
title: "Extraction of Data"
author: "Immanuel Williams PhD"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(rvest)
library(tidyverse)
library(httr)
library(dplyr)
library(purrr)
```



## Simplistic Call
```{r}
## Createtion of URL
team_name <- "was"
year <- 2023
generic_url <- paste0("https://www.pro-football-reference.com/teams/",team_name,"/",year,".htm#all_games")
```


```{r}
## Special Note 1: Maybe Optional (Further Research is needed on its relevance)
# request_details <- "Mozilla/5.0 (Windows NT 10.0; Win64; x64) AppleWebKit/537.36 (KHTML, like Gecko) Chrome/91.0.4472.124 Safari/537.36"
# response <- GET(url, user_agent(request_details),timeout(20))
```


```{r}
webpage <- generic_url %>% 
  rvest::read_html()
```


It is giving a memory address pointer instead of the actual data. 
[Can you go deeper on this]
> webpage[[1]]
<pointer: 0x0000022588f6e980>
Warning messages:
1: In xml_child(`__OBJECT__`, 1) : could not find function "xml_child"
2: In xml_child(`__OBJECT__`, 2) : could not find function "xml_child"
> webpage[[2]]
<pointer: 0x0000022584da4070>
> 

From the webpage, grab the HTML tables using rvest html_table()
```{r}
web_tables <- webpage %>% 
  rvest::html_table()
```
The result is a list containing HTML table elements.

Select the desired table, the 2023 Regular Season Table, which is the second table on the webpage. Use purr pluck() to select the table.
[Visit webpage to see available tables]
```{r}
int_web_table <- web_tables|> 
  purrr::pluck(2)
```

With our HTML table, use stats setNames to set the names of blank columns to "col". Using make.unique inside this function makes the names unique to avoid conflict, and the if_else function only applies this function to columns with the name "", or blank columns.
The "names()" function pulls the columns names into the function.

```{r}
int_web_table_temp <- int_web_table |> 
  janitor::clean_names()

new_colnames <- map_chr(seq_along(int_web_table_temp), ~ {
  part1 <- names(int_web_table_temp)[.x]
  part2 <- as.character(int_web_table_temp[1, .x])
  str_c(part1, part2, sep = "_")
})

web_table <- int_web_table_temp %>%
  slice(-c(1, 2)) %>%
  setNames(janitor::make_clean_names(new_colnames))

```


```{r}
web_table <- int_web_table |> 
  stats::setNames(make.unique(if_else(names(.) == "", "col", names(.))))
```

```{r}
web_table <- int_web_table |> 
  janitor::clean_names()
```

## Column Names Cleaning



Type 1: Columns with Both Header and First Row Values
- Example columns: 11-25 (Score, Offense, Defense columns)
- Initial state: Has both column header AND meaningful first row value
- Challenge: Need to combine both pieces of information with hyphen
- Solution needed: Merge header + first row value (e.g., "Score-Tm")
- Why special: Preserves complete context from both sources

Type 2: Empty Header and Empty First Row
- Example columns: 4-6, 9 (time, boxoffice, etc.)
- Initial state: No header AND no meaningful first row value
- Challenge: No useful information from either source
- Solution needed: Assign completely new custom names
- Why special: Requires external knowledge about what these columns represent

Type 3: Empty Header but Valuable First Row
- Example columns: 1-3, 7, 8, 10 (Week, Day, Date, etc.)
- Initial state: No header BUT has meaningful first row value
- Challenge: Only first row contains useful naming information
- Solution needed: Use first row value as column name
- Why special: Must preserve the meaningful information from first row only

Why One Size Doesn't Fit All:
1. Information Sources Vary:
   - Some columns have two sources of information
   - Some have no useful information
   - Some have partial information
   
2. Naming Requirements Differ:
   - Some need combined names
   - Some need completely new names
   - Some need preservation of existing values

3. Processing Order Matters:
   - Type 3 must be handled first to give empty columns initial names
   - Type 2 needs specific column positions after initial naming
   - Type 1 combines information last after other naming is stable

4. Error Handling Varies:
   - Type 1 needs checks for combining values
   - Type 2 needs precise position matching
   - Type 3 needs empty string checks


### Modificaiton 1 (Web Table): Type 3 Handling

- Creates first_row variable to store the first row values that we'll use for naming
- mod_1_table handles empty columns (Type 3)
- map2_chr compares each column name with first row values
- If column name starts with "col" (our temporary name), use the first row value
- Otherwise keep the original column name
- This affects columns 1-3, 7, 8, 10 which had empty names but values in first row

Use dplyr slice() to pull out the first row from our table, which contains headers for our table. Look at the table on the actual webpage to understand how this interacts with our current columns.
```{r}
first_row <- web_table %>% 
  dplyr::slice(1)
```



Use setNames again to integrate the "first row" column names with the current column headers. This will fill in names for columns titled with "col" that we established in the previous steps. To do this, we will also use purr map2_chr(), which maps ()
```{r}
mod_1_web_table <- web_table %>% 
  setNames(
    purrr::map2_chr(
      names(web_table),
      first_row,
      ~if(startsWith(.x, "col")){.y} else {.x}
    )
)
```


### Modification 2 (Web Table):  Type 2 Handling

- Handles the special case columns that need custom names
- Uses direct column position referencing with rename()
- Assigns custom names to columns 4 (Time), 5 (Boxoffice), 6 (Game Outcome), 9 (Game Location)
- These columns had empty names AND empty first row values

Now, we create a new modification to our table where we use dplyr rename() to set the name of certain columns. This is an instance of hard coding, where you code certain specific values instead of dynamic data like variables.  
```{r}
mod_2_web_table <- mod_1_web_table %>% 
  rename(
   "Time" = 4,
  #  "Boxoffice" = 5,
   "Game Outcome" = 8,
   "Game Location" = 6
 )
```


## Modificaiton 3 (Web Table): Type 1 Handling

- Processes columns 11-25 which have both header names and first row values
- match() finds the position of each column name
- For columns 11 and beyond:

1. Clean the original column name by removing any existing numbers and underscores
2. Clean the first row value by removing any existing numbers and underscores
3. Combine cleaned column name with cleaned first row value using underscore
4. Result example: "Score" + "Tm" = "score_tm" (instead of "score_1_tm")

- Keeps other column names unchanged (columns 1-10)
- slice(-1) removes the first row since its values are now part of column names
- Uses janitor::make_clean_names() for consistent naming format
- These columns require both pieces of information but need to avoid duplicate handling artifacts
- Prevents the creation of numbered suffixes that appear when R handles duplicate names

In this step, we will be incorporating the information in the first row to the current header.
```{r}
mod_3_web_table <- mod_2_web_table |>
  # Creating a new variable for our 3rd modification.
  setNames(
    janitor::make_clean_names(
      #wrap the code within this function to make names snakecase.
      purrr::map2_chr( # Map names to their correct columns
        names(mod_2_web_table),
        first_row, 
        # Pull out the names from the table, as well as the columns names of the first row.
        ~{
          col_index <- match(.x, names(mod_2_web_table))
          # loops through columns names and the corresponding values
          if(!is.na(col_index) && col_index >= 11) {
            #if the column index is 11 or higher 
            orig_name <- stringr::str_remove_all(.x, "[0-9_]")
            new_value <- stringr::str_remove_all(.y, "[0-9_]")
             #remove numbers and underscores
            paste(orig_name, new_value, sep = "_")
            #add underscore back, new name
          } else {
            .x
            #keep as is if not meeting the if condition
          }
        }
      )
    )
  ) |>
  dplyr::slice(-1)
#remove the first row, as we have now extracted all of the useful information from it.

```




We Should be Proud

Here we can use colnames() to see the resulting column names from this operation to ensure they are adequate.
```{r}
colnames(mod_3_web_table)
```

Next, we will use mutate() to convert some columns to numeric type. Using across(), we can apply the transformation to multiple columns at the same time
```{r}
mod_4_web_table <- mod_3_web_table %>% 
  mutate(
    # Convert numeric columns
    across(c(week, score_1, score_2, #changed from tm, opp
             starts_with("offense_"), 
             starts_with("defense_"),
             starts_with("expected_points_")), 
           as.numeric))
```

Now, we will use mutate() again to change the game_outcome type to a factor, and use case_when() to make more sense of the game_location column. By identifying when '@' appears, we can change it to away, and the opposite '' as home. We also change this to a factor. Lastly, we filter out the last column which is an aggregate column by filtering out games with no date.
```{r}

mod_5_web_table <- mod_4_web_table  %>% 

  mutate(
    game_outcome = as.factor(game_outcome),
    game_location = case_when(
      game_location == "@" ~ "away",
      game_location == "" ~ "home",
      TRUE ~ game_location
    ) %>% 
      as.factor()
  ) %>% 
  filter(day != "")
```















