---
title: "Visualization of Data"
author: "Immanuel Williams PhD"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(ggplot2)
library(dplyr)
library(forcats)
library(gato365dsh2024)
```


```{r}
load("was_historical_data.RData")
```

## Plot 1: Select Key Variable
### Gather Data
```{r}
plot_1_data_df <- was_historical_df %>% 
  select(score_tm) %>% 
  mutate(score_tm = as.numeric(score_tm))
```



### Plot Data
```{r}
plot_1_data_df %>% 
  ggplot(aes(x = score_tm)) +
  geom_density() +
  theme_custom()
```

Questions: 
Based the plot, describe the shape? [Statistics]
What elements of this need to be added to communcate the finding of this plot better? [Statistcs, Programming]
What statistic could be added to this plot to help display the center? [Statistcs, Programming]
What statistic could be added to this plot to help display the variability? [Statistcs, Programming]
What other visualizations could be used to display this information? Explain when you would want to use each type? [Statistics, Programming]



```{r}
plot_2_data_df <- was_historical_df %>% 
  select(score_tm,season)%>% 
  mutate(score_tm = as.numeric(score_tm),
         season = as.factor(season)
         )
```


```{r}
plot_2_data_df %>% 
ggplot(aes(x=season, y=score_tm)) +
  geom_violin() +
  theme_minimal() +
  labs(x="Season", y="Team Score", title = "Distribution of Team Scores by Season") +
  theme_custom()
```


Questions:
If you needed to create a table of this information what statics would be needed? [Statistics]
[General Questions on what they see?] [Statistics]
[If you already see the data, questions that are noticeable.] [stat]
What visualization could we add to this to see the 5 number sumary? [DS, stat, programming]
WHy would this be a good idea to add?













Main Question: How do team scores evolve across weeks of play within and between seasons?

```{r}
plot_3a_data_df <- was_historical_df %>% 
  select(score_tm,week,season) %>% 
  mutate(score_tm = as.numeric(score_tm),
         season = as.factor(season)
         ) %>% 
  
  filter(str_detect(week, "^[0-9]+$")) %>% 
  # mutate(week = fct_reorder(week, as.numeric(week))) %>% 
mutate(week_str = str_glue("w #{week}")) %>%
mutate(week_str = fct_reorder(week_str, as.numeric(week)))



plot_3b_data_df <- plot_3a_data_df %>% 
  group_by(week_str,season) %>% 
  summarise(`Mean tm Score` = mean(score_tm, na.rm = TRUE),
            `SD tm Score` = mean(score_tm, na.rm = TRUE),
            `Number of Games`= n(), 
            .groups = "drop"
            ) 
```

Questions:
Understand the observation unit well enough to know group_by when n() will be row for each? [DS]

Multiple Versions: 

```{r}

plot_3b_data_df %>% 
ggplot( aes(
  x = week_str, 
  y = `Mean tm Score`, 
   group = season, 
   color = season
  )) +
 geom_line() +
 geom_errorbar(aes(ymin = `Mean tm Score` - `SD tm Score`/sqrt(`Number of Games`),
                   ymax = `Mean tm Score` + `SD tm Score`/sqrt(`Number of Games`)),
               width = 0.2) +
 geom_point() +
 labs(title = "Mean Team Scores by Week and Season",
      y = "Mean Score",
      x = "Week") +
  theme_custom()
```


```{r}
plot_3b_data_df %>% 
ggplot( aes(
  x = season, 
  y = `Mean tm Score`, 
  color = week_str
  )) +
 geom_line() +
 geom_point() +
 labs(title = "Mean Team Scores by Week and Season",
      y = "Mean Score",
      x = "Week") +
  theme_custom()
```



Visualization Decision to show only filter years and week as the 

```{r}
selected_season <-  "2021"
plot_3b_data_df %>% 
  filter(season == selected_season) %>% 
  ggplot( aes(
  x = week_str, 
  y = `Mean tm Score`,
  group = 1
  )) +
 geom_line() +
 geom_point() +
 labs(title = str_glue("Mean Team Scores by Week \n for the {selected_season} Season"),
      y = "Mean Score",
      x = "Week") +
  theme_custom()
```



```{r}
selected_seasons <- 2019:2021
plot_3b_data_df %>% 
 filter(season %in% selected_seasons) %>% 
 ggplot(aes(
   x = week_str, 
   y = tm_score,
   color = season,
   group = season
 )) +
 geom_line() +
 geom_point() +
 labs(title = str_glue("Mean Team Scores by Week \n {min(selected_seasons)}-{max(selected_seasons)} Seasons"),
      y = "Mean Score",
      x = "Week") +
 theme_custom()
```









Questions:
Why is there two mutates? [DS, Programming]
Why Why am I still using tm rather than 'was'? [Programming, DS]
should the 1st mutate follow the second mutate? [DS, Programming]
Have the question that you are trying show, then have student reduce the number of variables or level or even change number of levels to make the plot more digestable? [DS, programming]
Let student start with code that is not ideal based on question and they are asked to fix it up to answer the question in the most fluid mechancism?


